%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% chapter1.tex
%% NOVA thesis document file
%%
%% Chapter with introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT LOADING FILE chapter1.tex}

\chapter{Introduction}
\label{cha:introduction}

% epigraph configuration
\epigraphfontsize{\small\itshape}
\setlength\epigraphwidth{12.5cm}
\setlength\epigraphrule{0pt}

\epigraph{
  This work is licensed under the \href{LaTeX project public license}{\LaTeX\ Project Public License v1.3c}.
  To view a copy of this license, visit \url{LaTeX project public license}.
}

\section{Context}

Consensus protocols have crucial roles among distributed systems, such as blockchain technologies. They enable coordination between multiple machines, allowing them to maintain a replicated state and are therefore crucial in building reliable, large-scale distributed systems.

Blockchain protocols are a composition of steps executed by each node in a network to maintain a distributed ledger among multiple nodes, and also enables the development of smart contracts.

These blockchain protocols possess several underlying components, or algorithms, and each of which may have several parameters that can influence the performance and behavior of the overall protocol. It is therefore important to be able to reason about the choice of these parameters before deploying the protocol to a real environment.

In recent years, blockchain technologies have been rising in popularity leading to the emergence of several different protocols, algorithms and parameterizations. Hence, there is a need to divide blockchain protocols into essential components, to achieve better abstraction between different protocols, and extract their similarities and key differences. These components can then be used as building blocks to structure the presentation and implementation of the protocols, enhancing their readability and code reusability, and thus enabling the development of an extensible simulation environment.

Chapter \ref{cha:background} will present consensus and blockchain protocols in further detail.


\section{Problem}

Consensus protocols are notoriously difficult to understand and even harder to define correctly. In the context of blockchain systems, mastering their dynamical behaviour and trust in their correctness is crucial, since the protocols control financial transactions. Moreover, understanding them is essential to enable evolution. However,

\begin{center}
  \emph{there is a lack of specification and experimentation environments to develop consensus and blockchain protocols and thus study their behaviour}
\end{center}

\noindent which is the problem this thesis aims to address.



\section{Goal}

The goal of this thesis is to develop a simulation environment that allows researchers to make informed decisions about the effects of different parameterizations of the underlying components involved in blockchain protocols.

The simulation environment should be:

\begin{itemize}
  \item modular and extensible, providing the ability to make changes to the protocols, and simulate different families of protocols.
  \item parameterizable, to allow researchers to learn the effects of different parameters in the overall behaviour of the protocol.
  \item structured according to an abstraction framework, such as the five component framework presented in \textit{"Survey Of Distributed Consensus Protocols for Blockchain Networks"} \cite{survey_bchain_networks}, enabling modularity, code reusability and readability.
  \item focused on providing a qualitative evaluation of the simulated protocols.
\end{itemize}

The goal is \textbf{not} to simulate the protocol's behaviour when compared to reality as a mean fo testing it's efficiency, but rather to provide researchers results that allow them to reason about certain properties of the implementations, such as: How many adversaries does the system support? Is the system fair when handling stakeholders with different stakes?


This thesis is being developed in the context of NOVA LINCS research unit, in parallel with two other projects which aim to supply researchers with tools that facilitate the empirical analysis of blockchain protocols.

The other projects consist in producing a DSL (Domain Specific Language) for specifying these protocols, which could then be compiled and executed in the simulator and also compiled to another environment where formal proofing will be executed.

